---
title: "Exercise-12"
author: "Sarah"
date: "2024-04-10"
output: html_document
---
Load Libraries
```{r, eval=TRUE, echo=TRUE, warning=FALSE}
library(tidyverse)
library(skimr)
library(MASS)
library(MuMIn)
```
Load in data
```{R}
#Ernest 2003. Life history characteristics of placental mammals
f<-"https://raw.githubusercontent.com/difiore/ada-2024-datasets/main/Mammal_lifehistories_v2.txt"
d<-read_tsv(f, col_names = TRUE)
head(d)

skim(d)

#All variables are numeric except for taxonomic info.
```
1. Replace -999 with NA and reskim
```{R}
#Replace 999 with NA
d[d == -999.00] <- NA

#Reskim
skim(d)
```
2. Drop litter size and refs
```{r}
d<- subset(d, select = -c(`litter size`, `refs`))
head(d)
```
3. Log transform all numeric variables
```{R}
d<-d%>%mutate(logMass=log(`mass(g)`),
              logGest=log(`gestation(mo)`),
              lognb=log(`newborn(g)`),
              logwean=log(`weaning(mo)`),
              logwm=log(`wean mass(g)`),
              logAFR=log(`AFR(mo)`),
              logml=log(`max. life(mo)`),
              logly=log(`litters/year`))
head(d)
```
4. Regress log transformed age and mass variables on overall mass. Add residuals to dataframe.
```{R}
#Regression
m1<-lm(data=d, logGest~logMass, na.action = na.exclude)
m2<-lm(data=d, logwean~logMass, na.action = na.exclude)
m3<-lm(data=d, logAFR~logMass, na.action = na.exclude)
m4<-lm(data=d, logml~logMass, na.action = na.exclude)
m5<-lm(data=d, lognb~logMass, na.action = na.exclude)
m6<-lm(data=d, logwm~logMass, na.action = na.exclude)

#Add residuals to dataframe
d$relGest<-residuals(m1)
d$relWean<-residuals(m2)
d$relAFR<-residuals(m3)
d$relLife<-residuals(m4)
d$relNM<-residuals(m5)
d$relWM<-residuals(m6)
```
5. Make plots with residuals. 
```{R}
#Max lifespan in relation to order
a<-ggplot(d, aes(y=relLife, x=order))+
  geom_boxplot()+geom_jitter(width=0.2)+
  theme(axis.text.x = element_text(angle = 45)) 
a

#Newborn mass in relation to Order
b<-ggplot(d, aes(y=relNM, x=order))+
  geom_boxplot()+geom_jitter(width=0.2)+
  theme(axis.text.x = element_text(angle = 45)) 
b

#Weaning mass in relation to Order
c<-ggplot(d, aes(y=relWM, x=order))+
  geom_boxplot()+geom_jitter(width=0.2)+
  theme(axis.text.x = element_text(angle = 45)) 
c

```

Primates have the highest residual lifespan. Macroscelididae (and a couple rodent species) have the highest residual newborn mass. Perissodactyla has the highest residual weaning mass. 

6a. Model selection for two response variables
```{R}
#Remove NAs
d<-d%>%drop_na(logml, logAFR, logGest, lognb, logwean, logwm, logly, logMass)


#Model selection for log(max.life)
ml_null<-lm(data=d, logml~1)

#Use MASS package
s<-stepAIC(ml_null, scope=.~.+logGest+lognb+logwean+logwm+logly+logMass, direction = "both", trace = TRUE)
#add everything except lognb and logwm to achieve smallest (most negative) AIC. 2 other models have delta AICs<4. 

ml<-lm(data=d, logml~logGest+logly+logMass+logwean)
summary(ml)


#Model selection for log(AFR)
mAFR_null<-lm(data=d, logAFR~1)

#Use MASS package
t<-stepAIC(mAFR_null, scope=.~.+logGest+lognb+logwean+logwm+logly+logMass, direction = "both", trace = TRUE)
#add everything except lognb and logwm to achieve smallest (most negative) AIC. 4 other models have delta AICs<4. 

mAFR<-lm(data=d, logAFR ~ logGest+logly+logMass+logwean)
summary(mAFR)


#Both models include logly, logGest, logwean, and logMass as predictor variables. 
```
6b. Calculate and plot the model-averaged coefficients and their CIs across this set of top models.
```{R}
#Log(max.Life)

#full model
ml_full<-lm(data=d, logml~logGest+lognb+logwean+logwm+logly+logMass, na.action = na.fail)

ml.set <- dredge(ml_full, beta = "none", m.lim = c(0, 6))

#Avg coeffs across models
(m.avg.l <- summary(model.avg(ml.set, subset = delta < 4, fit = TRUE)))

confint(m.avg.l)

#Plot
plot(m.avg.l, full = TRUE, intercept = FALSE)


#Log(AFR)

#full model
mAFR_full<-lm(data=d, logml~logGest+lognb+logwean+logwm+logly+logMass, na.action = na.fail)

mAFR.set <- dredge(mAFR_full, beta = "none", m.lim = c(0, 6))

#Avg coeffs across models
(m.avg.AFR <- summary(model.avg(mAFR.set, subset = delta < 4, fit = TRUE)))

confint(m.avg.AFR)

#Plot
plot(m.avg.AFR, full = TRUE, intercept = FALSE)

```

logly, logGest, logwean, and logMass are best predictors for log(Max.Life) and Log(AFR). 







